# Chronology & Age Related Tables

## AgeTypes {#AgeTypes}

Lookup table of Age Types or units. This table is referenced by the [`chronologies`](#Chronologies) and [`geochronology`](#Geochronology) tables.

```{r, agetypestable, echo=FALSE}
showTable(db, 'agetypes')
```

* **AgeTypeID (Primary Key)**: An arbitrary Age Type identification number.
* **AgeType**: Age type or units. Includes the following:
  * Calendar years AD/BC
  * Calendar years BP
  * Calibrated radiocarbon years BP
  * Radiocarbon years BP
  * Varve years BP

## AggregateChronologies {#AggregateChronologies}

This table stores metadata for Aggregate Chronologies. An Aggregate Chronology refers to an explicit chronology assigned to a sample Aggregate. The individual Aggregate Samples have ages assigned in the [`aggregatesampleages`](#AggregateSampleAges) table. An Aggregate Chronology would be used, for example, for a set of packrat middens assigned to an [`aggregatedatasets`](#AggregateDatasets). The Aggregate Chronology is analagous to the Chronology assigned to samples from a single Collection Unit.

An Aggregate may have more than one Aggregate Chronology, for example one in radiocarbon years and another in calibrated radiocarbon years. One Aggreagate Chronology per Age Type may be designated the default, which is the Aggregate Chronology currently preferred by the database stewards.

```{r, AggregateChronologiestable, echo=FALSE}
showTable(db, 'aggregatechronologies')
```

* **AggregateChronID (Primary Key)**: An arbitrary Aggregate Chronology identification number.
* **AggregateDatasetID (Foreign Key)**: Dataset to which the Aggregate Chronology applies. Field links to the [`aggregatedatasets`](#AggregateDatasets) table.
* **AgeTypeID (Foreign Key)**: Age type or units. Field links to the [`agetypes`](#AgeTypes) table.
* **IsDefault**: Indicates whether the Aggregate Chronology is a default or not. Default status is determined by a Neotoma data steward. Aggregate Datasets may have more than one default Aggregate Chronology, but may have only one default Aggregate Chronology per Age Type.
* **ChronologyName**: Optional name for the Chronology.
* **AgeBoundYounger**: The younger reliable age bound for the Aggregate Chronology. Younger ages may be assigned to samples, but are not regarded as reliable. If the entire Chronology is considered reliable, AgeBoundYounger is assigned the youngest sample age rounded down to the nearest 10. Thus, for 72 BP, AgeBoundYounger = 70 BP; for -45 BP, AgeBoundYounger = -50 BP.
* **AgeBoundOlder**: The older reliable age bound for the Aggregate Chronology. Ages older than AgeOlderBound may be assigned to samples, but are not regarded as reliable. This situation is particularly true for ages extrapolated beyond the oldest Chron Control. . If the entire Chronology is considered reliable, AgeBoundOlder is assigned the oldest sample age rounded up to the nearest 10. Thus, for 12564 BP, AgeBoundOlder is 12570.
* **Notes**: Free form notes or comments about the Aggregate Chronology.

## ChronControls {#ChronControls}

This table stores data for Chronology Controls, which are the age-depth control points used for age models. These controls may be geophysical controls, such as radiocarbon dates, but include many other kinds of age controls, such as biostratigraphic controls, archaeological cultural associations, and volcanic tephras. In the case of radiocarbon dates, a Chronology Control may not simply be the raw radiocarbon date reported by the laboratory, but perhaps a radiocarbon date corrected for an old carbon reservoir, a calibrated radiocarbon date, or an average of several radiocarbon dates from the same level. A common control for
lake-sediment cores is the age of the top of the core, which may be the year the core was taken or perhaps an estimate of 0 BP if a few cm of surficial sediment were lost.

```{r, ChronControlstable, echo=FALSE}
showTable(db, 'chroncontrols')
```

* **ChronControlID (Primary Key)**: An arbitrary Chronology Control identification number.
* **ChronologyID (Foreign Key)**: Chronology to which the ChronControl belongs. Field links to the Chronolgies table.
* **ChronControlTypeID (Foreign Key)**: The type of Chronology Control. Field links to the [`chroncontroltypes`](#ChronControlTypes) table.
* **Depth**: Depth of the Chronology Control in cm.
* **Thickness**: Thickness of the Chronology Control in cm.
* **Age**: Age of the Chronology Control.
* **AgeLimitYounger**: The younger age limit of a Chronology Control. This limit may be explicitly defined, for example the younger of the 2-sigma range limits of a calibrated radiocarbon date, or it may be more loosely defined, for example the younger limit on the range of dates for a biostratigraphic horizon.
* **AgeLimitOlder**: The older age limit of a Chronology Control.
* **Notes**: Free form notes or comments about the Chronology Control.

## ChronControlTypes {#ChronControlTypes}

Lookup table of Chronology Control Types. This table is referenced by the [`chroncontrols`](#ChronControls) table.

```{r, ChronControlTypestable, echo=FALSE}
showTable(db, 'chroncontroltypes')
```

**ChronControlTypeID (Primary Key)**

:   An arbitrary Chronology Control Type identification number.

**ChronControlType**

:   The type of Chronology Control object. Chronology Controls include
    such geophysical controls as radiocarbon dates, calibrated
    radiocarbon dates, averages of several radiocarbon dates,
    potassium-argon dates, and thermoluminescence dates, as well as
    biostratigraphic controls, sediment stratigraphic contols, volcanic
    tephras, archaeological cultural associations, and any other types
    of age controls. In general these are calibrated or calendar year
    dates Before Present (BP). **Some ChronControlTypes are in Radiocarbon
    Years, so caution must be exercised.**

## Chronologies {#Chronologies}

This table stores Chronology data. A Chronology refers to an explicit
chronology assigned to a Collection Unit. A Chronology has Chronology
Controls, the actual age-depth control points, which are stored in the
[`chroncontrols`](#ChronControls) table. A Chronology is also based on an Age
Model, which may be a numerical method that fits a curve to a set of
age-depth control points or may simply be individually dated Analysis
Units.

A Collection Unit may have more than one Chronology, for example one in
radiocarbon years and another in calibrated radiocarbon years. There may
be a Chronology developed by the original author and another developed
by a later research project. Chronologies may be stored for archival
reasons, even though they are now believed to have problems, if they
were used for an important research project. One Chronology per Age Type
may be designated the default Chronology, which is the Chronology
currently preferred by the database stewards.

Based upon the Chronology, which includes the Age Model and the Chron
Controls, ages are assigned to individual samples, which are stored in
the [`SampleAges`](#SampleAges) table.

A younger and older age bounds are assigned to the Chronology. Within
these bounds the Chronology is regarded as reliable. Ages may be
assigned to samples beyond the reliable age bounds, but these are not
considered reliable.

```{r, Chronologiestable, echo=FALSE}
showTable(db, 'chronologies')
```

**ChronologyID (Primary Key)**

:   An arbitrary Chronology identification number.

**CollectionUnitID (Foreign Key)**

:   Collection Unit to which the Chronology applies. Field links to the
    `CollectionUnits`{.interpreted-text role="ref"} table.

**AgeTypeID (Foreign Key)**

:   Age type or units. Field links to the `AgeTypes`{.interpreted-text
    role="ref"} table.

**ContactID (Foreign Key)**

:   Person who developed the Age Model. Field links to the
    `Contacts`{.interpreted-text role="ref"} table.

**IsDefault**

:   Indicates whether the Chronology is a default chronology or not.
    Default status is determined by a Neotoma data steward. Collection
    Units may have more than one default Chronology, but may have only
    one default Chronology per Age Type. Thus, there may be a default
    radiocarbon year Chronology and a default calibrated radiocarbon
    year Chronology, but only one of each. Default Chronologies may be
    used by the Neotoma web site, or other web sites, for displaying
    default diagrams or time series of data. Default Chronologies may
    also be of considerable use for actual research purposes; however,
    users may of course choose to develop their own chronologies.

**ChronologyName**

:   Optional name for the Chronology. Some examples are:

    -   COHMAP chron 1 A Chronology assigned by the COHMAP project.
    -   COHMAP chron 2 An alternative Chronology assigned by the COHMAP
        project
    -   NAPD 1 A Chronology assigned by the North American Pollen
        Database.
    -   Gajewski 1995 A Chronology assigned by Gajewski (1995).

**DatePrepared**

:   Date that the Chronology was prepared.

**AgeModel**

:   The age model used for the Chronology. Some examples are: linear
    interpolation, 3^rd^ order polynomial, and individually dated
    analysis units.

**AgeBoundYounger**

:   The younger reliable age bound for the Chronology. Younger ages may
    be assigned to samples, but are not regarded as reliable. If the
    entire Chronology is considered reliable, AgeBoundYounger is
    assigned the youngest sample age rounded down to the nearest 10.
    Thus, for 72 BP, AgeBoundYounger = 70 BP; for -45 BP,
    AgeBoundYounger = -50 BP.

**AgeBoundOlder**

:   The older reliable age bound for the Chronology. Ages older than
    AgeOlderBound may be assigned to samples, but are not regarded as
    reliable. This situation is particularly true for ages extrapolated
    beyond the oldest Chron Control. . If the entire Chronology is
    considered reliable, AgeBoundOlder is assigned the oldest sample age
    rounded up to the nearest 10. Thus, for 12564 BP, AgeBoundOlder
    is 12570.

**Notes**

:   Free form notes or comments about the Chronology.

### SQL Example

The following SQL statement produces a list of Chronologies for a particular site. In this case we are matching on the site name, for "Irvin Lake", :

```{sql, connection=db}
SELECT st.sitename,
       ch.chronologyid,
       ch.chronologyname,
       ch.isdefault,
       at.agetype
FROM ndb.sites AS st
INNER JOIN ndb.collectionunits AS cu ON cu.siteid = st.siteid
INNER JOIN ndb.chronologies AS ch ON ch.collectionunitid = cu.collectionunitid
INNER JOIN ndb.agetypes AS at ON at.agetypeid = ch.agetypeid
WHERE st.sitename = 'Irvin Lake';
```

From this we see that there have been five chronologies generated for Irvin Lake. There are two [age types](#AgeTypes) represented, *Radiocarbon years* and *Calibrated radiocarbon years BP*.  There is one default age model for *Calibrated radiocarbon years BP* and, at the time of this writing, the `NAPD 1` data model is not defined as the default. This is counter to the general data entry standard, where each age type has a defined default chronology.

### SQL Example

The following statement produces a list of the [chronological controls](#ChronControls) for the default chronology from **Wolsfeld Lake** in Calibrated radiocarbon years BP. Here we show only the first 5 rows:

```{sql, connection=db}
SELECT cc.depth, ChronControls.Age,
ChronControls.AgeLimitYounger, ChronControls.AgeLimitOlder,
ChronControlTypes.ChronControlType
FROM ndb.sites AS st
INNER JOIN ndb.collectionunits AS cu ON cu.siteid = st.siteid
INNER JOIN ndb.chronologies AS ch ON ch.collectionunitid = cu.collectionunitid
INNER JOIN ndb.agetypes ON 
dd
FROM ChronControlTypes cINNER JOIN ((AgeTypes INNER JOIN ((Sites INNER
JOIN CollectionUnits ON Sites.SiteID = CollectionUnits.SiteID) INNER
JOIN Chronologies ON CollectionUnits.CollectionUnitID =
Chronologies.CollectionUnitID) ON AgeTypes.AgeTypeId =
Chronologies.AgeTypeID) INNER JOIN ChronControls ON
Chronologies.ChronologyID = ChronControls.ChronologyID) ON
ChronControlTypes.ChronControlTypeID = ChronControls.ChronControlTypeID

WHERE (((Sites.SiteName)="Wolsfeld Lake") AND
((Chronologies.IsDefault)=True) AND ((AgeTypes.AgeType)="Calibrated
radiocarbon years BP"));
```

Result:

  ----------- --------- --------------------- ------------------- --------------------------
  **Depth**   **Age**   **AgeLimitYounger**   **AgeLimitOlder**   **ChronControlType**

  650         -25       -25                   -25                 Core top

  662         -13       -8                    -18                 Interpolated, corrected
                                                                  for compaction

  670         0         -5                    5                   Interpolated, corrected
                                                                  for compaction

  680         22        17                    27                  Interpolated, corrected
                                                                  for compaction

  690         46        41                    51                  Interpolated, corrected
                                                                  for compaction

  . . .       . . .     . . .                 . . .               . . .
  ----------- --------- --------------------- ------------------- --------------------------

## AggregateSampleAges {#AggregateSampleAges}

This table stores the links to the ages of samples in an Aggregate
Dataset. The table is necessary because samples may be from Collection
Units with multiple chronologies, and this table stores the links to the
sample ages desired for the Aggregate Dataset.

```{r, aggregatesampleagestable, echo=FALSE}
showTable(db, 'aggregatesampleages')
```

**AggregateDatasetID (Primary Key, Foreign Key)**

:   Aggregate Dataset identification number. Field links to the
    `AggregateDatasets`{.interpreted-text role="ref"} table.

**AggregateChronID (Primary Key, Foreign Key)**

:   Aggregate Chronology identification number Field links to the
    `AggregateChronologies`{.interpreted-text role="ref"} table.

**SampleAgeID (Primary Key, Foreign Key)**

:   Sample Age ID number. Field links to the
    `SampleAges`{.interpreted-text role="ref"} table.

### SQL Example

The following SQL statement produces a list of Sample ID numbers and
ages for the Aggregate Dataset:

``` {.sourceCode .sql}
SELECT AggregateSamples.SampleID, SampleAges.Age

FROM SampleAges INNER JOIN ((AggregateDatasets INNER JOIN
AggregateSampleAges ON AggregateDatasets.AggregateDatasetID =
AggregateSampleAges.AggregateDatasetID) INNER JOIN AggregateSamples ON
AggregateDatasets.AggregateDatasetID =
AggregateSamples.AggregateDatasetID) ON (AggregateSamples.SampleID =
SampleAges.SampleID) AND (SampleAges.SampleAgeID =
AggregateSampleAges.SampleAgeID)

WHERE (((AggregateDatasets.AggregateDatasetName)=""));
```

### SQL Example

The AggregateSampleAges table may have multiple SampleAgeID\'s for
Aggregate Dataset samples, for example SampleAgeID\'s for radiocarbon
and calibrated radiocarbon chronologies. In this case, the Chronolgies
table must be linked into a query to obtain the ages of Aggregate
Samples, and either the AgeTypeID must be specified in the Chronolgies
table or the `AgeTypes`{.interpreted-text role="ref"} table must also be
linked with the AgeType specified. The following SQL statement produces
a list of Sample ID numbers and «Radiocarbon years BP» ages for the «»
Aggregate Dataset: `Samples`{.interpreted-text role="ref"}

``` {.sourceCode .sql}
SELECT AggregateSamples.SampleID, SampleAges.Age

FROM AgeTypes INNER JOIN (Chronologies INNER JOIN (SampleAges INNER JOIN
((AggregateDatasets INNER JOIN AggregateSampleAges ON
AggregateDatasets.AggregateDatasetID =
AggregateSampleAges.AggregateDatasetID) INNER JOIN AggregateSamples ON
AggregateDatasets.AggregateDatasetID =
AggregateSamples.AggregateDatasetID) ON (AggregateSamples.SampleID =
SampleAges.SampleID) AND (SampleAges.SampleAgeID =
AggregateSampleAges.SampleAgeID)) ON Chronologies.ChronologyID =
SampleAges.ChronologyID) ON AgeTypes.AgeTypeId = Chronologies.AgeTypeID

WHERE (((AggregateDatasets.AggregateDatasetName)="") AND
((AgeTypes.AgeType)="Radiocarbon years BP"));
```

## Geochronology {#Geochronology}

This table stores geochronologic data. Geochronologic measurements are
from geochronologic samples, which are from Analysis Units, which may
have a depth and thickness. Geochronologic measurments may be from the
same or different Analysis Units as fossils. In the case of faunal
excavations, geochronologic samples are typically from the same Analysis
Units as the fossils, and there may be multiple geochronologic samples
from a single Analysis Unit. In the case of cores used for microfossil
analyses, geochronologic samples are often from separate Analysis Units;
dated core sections are often thicker than microfossil Analysis Units.

```{r, geochronologytable, echo=FALSE}
showTable(db, 'geochronology')
```

**GeochronID (Primary Key)**

:   An arbitrary Geochronologic identificantion number.

**SampleID (Foreign Key)**

:   Sample identification number. Field links to
    `Samples`{.interpreted-text role="ref"} table.

**GeochronTypeID (Foreign Key)**

:   Identification number for the type of Geochronologic analysis, e.g.
    «Carbon-14», «Thermoluminescence». Field links to the
    `GeochronTypes`{.interpreted-text role="ref"} table.

**AgeTypeID (Foreign Key)**

:   Identification number for the age units, e.g. «Radiocarbon years
    BP», «Calibrated radiocarbon years BP».

**Age**

:   Reported age value of the geochronologic measurement.

**ErrorOlder**

:   The older error limit of the age value. For a date reported with ±1
    SD or σ, the ErrorOlder and ErrorYounger values are this value.

**ErrorYounger**

:   The younger error limit of the age value.

**Infinite**

:   Is «True» for and infinite or "greater than" geochronologic
    measurement, otherwise is «False».

**Delta13C**

:   The measured or assumed δ^13^C value for radiocarbon dates, if
    provided. Radiocarbon dates are assumed to be normalized to δ^13^C,
    and if uncorrected and normalized ages are reported, the normalized
    age should be entered in the database.

**LabNumber**

:   Lab number for the geochronologic measurement.

**Material Dated**

:   Material analyzed for a geochronologic measurement.

**Notes**

:   Free form notes or comments about the geochronologic measurement.

### SQL Example

This query lists the geochronologic data for Montezuma Well.

``` {.sourceCode .sql}
SELECT AnalysisUnits.Depth, AnalysisUnits.Thickness, 
GeochronTypes.GeochronType, Geochronology.Age, Geochronology.ErrorOlder,
Geochronology.ErrorYounger, Geochronology.Delta13C,
Geochronology.LabNumber, Geochronology.MaterialDated,
Geochronology.Notes

FROM GeochronTypes INNER JOIN ((((Sites INNER JOIN CollectionUnits ON
Sites.SiteID = CollectionUnits.SiteID) INNER JOIN AnalysisUnits ON
CollectionUnits.CollectionUnitID = AnalysisUnits.CollectionUnitID) INNER
JOIN Samples ON AnalysisUnits.AnalysisUnitID = Samples.AnalysisUnitID)
INNER JOIN Geochronology ON Samples.SampleID = Geochronology.SampleID)
ON GeochronTypes.GeochronTypeID = Geochronology.GeochronTypeID

WHERE (((Sites.SiteName)="Montezuma Well"));
```

Result:

  ----------- ------------- ------------------ --------- --------- ----------- -------------- ---------- ------------------- --------------
  **Depth**   **Thick..**   **GeochronType**   **Age**   **Error   **Error     **Delta13C**   **Lab      **MaterialDated**   **Notes**
                                                         Older**   Younger**                  Number**                       

  1015        1             Carbon-14:         10975     95        95                         AA-4694    Juniperus twig       
                            accelerator mass                                                                                 
                            spectrometry                                                                                     

  225         10            Carbon-14:         1526      50        50                         AA-2450    charcoal, wood       
                            accelerator mass                                                                                 
                            spectrometry                                                                                     

  330         10            Carbon-14:         2885      60        60                         AA-2451    charcoal, wood       
                            accelerator mass                                                                                 
                            spectrometry                                                                                     

  395         10            Carbon-14:         5540      60        60                         AA-4693    charcoal, wood       
                            accelerator mass                                                                                 
                            spectrometry                                                                                     

  465         10            Carbon-14:         8003      70        70                         AA-2452    Scirpus achenes      
                            accelerator mass                                                                                 
                            spectrometry                                                                                     

  535         10            Carbon-14:         14950     350       320         -26.7          A-4732     bark                Davis and
                            proportional gas                                                                                 Shafer (1992)
                            counting                                                                                         reject as too
                                                                                                                             old.

  887         1             Carbon-14:         9520      200       200         -25.3          A-4733     wood                 
                            proportional gas                                                                                 
                            counting                                                                                         

  887         1             Carbon-14:         24910     370       370                        AA-5053    wood                Davis and
                            accelerator mass                                                                                 Shafer (1992)
                            spectrometry                                                                                     reject as too
                                                                                                                             old.
  ----------- ------------- ------------------ --------- --------- ----------- -------------- ---------- ------------------- --------------

## GeochronPublications {#GeochronPublications}

Publications in which Geochronologic measurements are reported. Many
older radiocarbon dates are reported in the journal *Radiocarbon*. Dates
may be reported in multiple publications. The \"publication\" could be a
database such as the online Canadian Archaeological Radiocarbon
Database.

```{r, geochronpublicationstable, echo=FALSE}
showTable(db, 'geochronpublications')
```

**GeochronID (Primary Key, Foreign Key)**

:   Geochronologic identification number. Field links to the
    `Geochronology`{.interpreted-text role="ref"} table.

**PublicationID (Primary Key, Foreign Key)**

:   Publication identification number. Field links to the
    `Publications`{.interpreted-text role="ref"} table.

## GeochronTypes {#GeochronTypes}

Lookup table for Geochronology Types. Table is referenced by the
`Geochronology`{.interpreted-text role="ref"} table.

  ----------------------------------------------------
  Field Name                   Variable Type    Key
  ---------------------------- ---------------- ------
  GeochronTypeID               Long Integer     PK

  GeochronType                 Text             
  ----------------------------------------------------

**GeochronTypeID (Primary Key)**

:   Geochronology Type identification number.

**GeochronType**

:   Type of Geochronologic measurement.

## RelativeAgePublications {#RelativeAgePublications}

This table stores Publications in which Relative Ages are reported for
CollectionUnits.

  ----------------------------------------------------------------------------------------
  Field Name                   Variable Type  Key       Reference Table
  ---------------------------- -------------- --------- ----------------------------------
  RelativeAgeID                Long Integer   PK, FK    `RelativeAges`{.interpreted-text
                                                        role="ref"}

  PublicationID                Long Integer   PK, FK    `Publications`{.interpreted-text
                                                        role="ref"}
  ----------------------------------------------------------------------------------------

**RelativeAgeID (Primary Key, Foreign Key)**

:   Relative Ages identification number. Field links to the
    `RelativeAges`{.interpreted-text role="ref"} table.

**PublicationID (Primary Key, Foreign Key)**

:   Publication identification number. Field links to
    `Publications`{.interpreted-text role="ref"} table.

## RelativeAges {#RelativeAges}

Lookup table of RelativeAges. Table is referenced by the
`RelativeChronology`{.interpreted-text role="ref"} table.

  --------------------------------------------------------------------------------------
  Field Name               Variable Type  Key    Reference Table
  ------------------------ -------------- ------ ---------------------------------------
  RelativeAgeID            Long Integer   PK      

  RelativeAgeUnitID        Long Integer   FK     `RelativeAgeUnits`{.interpreted-text
                                                 role="ref"}

  RelativeAgeScaleID       Long Integer   FK     `RelativeAgeScales`{.interpreted-text
                                                 role="ref"}

  RelativeAge              Text                   

  C14AgeYounger            Double                 

  C14AgeOlder              Double                 

  CalAgeYounger            Double                 

  CalAgeOlder              Double                 

  Notes                    Memo                   
  --------------------------------------------------------------------------------------

**RelativeAgeID (Primary Key)**

:   An arbitrary Relative Age identification number.

**RelativeAgeUnitID (Foreign Key)**

:   Relative Age Unit (e.g. «Marine isotope stage», «Land mammal age»).
    Field links to the `RelativeAgeUnits`{.interpreted-text role="ref"}
    lookup table.

**RelativeAgeScaleID (Foreign Key)**

:   Relative Age Scale (e.g. «Geologic time scale», «Marine isotope
    stages»). Field links to the `RelativeAgeScales`{.interpreted-text
    role="ref"} lookup table.

**RelativeAge**

:   Relative Age (e.g. «Rancholabrean», a land mammal age; «MIS 11»,
    marine isotope stage 11).

**C14AgeYounger**

:   Younger age of the Relative Age unit in ^14^C yr B.P. Applies only
    to Relative Age units within the radiocarbon time scale.

**C14AgeOlder**

:   Older age of the Relative Age unit in ^14^C yr B.P. Applies only to
    Relative Age units within the radiocarbon time scale.

**CalAgeYounger**

:   Younger age of the Relative Age unit in calendar years.

**CalAgeOlder**

:   Older age of the Relative age unit in calendar years.

**Notes**

:   Free form notes or comments about Relative Age unit.

SQL Example
-----------

The following query gives the Relative Ages for the «North American land
mammal ages». The Relative Age Unit for each of these is «Land mammal
age». Commas were added to the ages in the query result to make them
more readable.

``` {.sourceCode .sql}
SELECT RelativeAges.RelativeAge, RelativeAges.CalAgeYounger,
RelativeAges.CalAgeOlder

FROM RelativeAgeScales INNER JOIN RelativeAges ON
RelativeAgeScales.RelativeAgeScaleID = RelativeAges.RelativeAgeScaleID

WHERE (((RelativeAgeScales.RelativeAgeScale)="North American land mammal ages"));
```

Result:

  CalAgeYounger   CalAgeOlder   RelativeAge
  --------------- ------------- -------------------
  14000.0         150000.0      Rancholabrean
  150000.0        1900000.0     Irvingtonian
  850000.0        1900000.0     Irvingtonian I
  400000.0        850000.0      Irvingtonian II
  150000.0        400000.0      Irvingtonian III
  1900000.0       4900000.0     Blancan
  4620000.0       4900000.0     Blancan I
  4100000.0       4620000.0     Blancan II
  3000000.0       4100000.0     Blancan III
  2500000.0       3000000.0     Blancan IV
  1900000.0       2500000.0     Blancan V
  4900000.0       9000000.0     Hemphillian
  10000.0         14000.0       Early Santarosean
  410.0           10000.0       Late Santarosean
  -100.0          410.0         Saintaugustinean
  410.0           14000.0       Santarosean
                                

RadiocarbonCalibration {#RadiocarbonCalibration}
======================

Radiocarbon calibraton table. This table is intended for quick
calibraton of age-model radiocarbon dates. These calibrated dates are
for perusal and data exploration only. Please see Section *2.5* for a
full discussion.

  -------------------------------------------------------------
  Field Name                            Variable Type    Key
  ------------------------------------- ---------------- ------
  C14yrBP                               Long Integer     PK

  CalyrBP                               Long Integer     
  -------------------------------------------------------------

**C14yrBP**

:   Age in radiocarbon years BP. The range is -100 to 45,000 by 1-year
    increments.

**CalyrBP**

:   Age in calibrated radiocarbon years BP.

RelativeAgeScales {#RelativeAgeScales}
=================

Lookup table of Relative Age Scales. Table is referenced by the
`RelativeAges`{.interpreted-text role="ref"} table.

  --------------------------------------------------------
  Field Name                       Variable Type    Key
  -------------------------------- ---------------- ------
  RelativeAgeScaleID               Long Integer     PK

  RelativeAgeScale                 Text             
  --------------------------------------------------------

**RelativeAgeScaleID (Primary Key)**

:   An arbitrary Relative Age Scale identification number.

**RelativeAgeScale**

:   Relative Age Scale. The table stores the following Relative Age
    Scales:

    -   Archaeological time scale
    -   Geologic time scale
    -   Geomagnetic polarity time scale
    -   Marine isotope stages
    -   North American land mammal ages
    -   Quaternary event classification

RelativeAgeUnits {#RelativeAgeUnits}
================

Lookup table of RelativeAgeUnits. Table is referenced by the
`RelativeAges`{.interpreted-text role="ref"} table.

  -------------------------------------------------------
  Field Name                      Variable Type    Key
  ------------------------------- ---------------- ------
  RelativeAgeUnitID               Long Integer     PK

  RelativeAgeUnit                 Text             
  -------------------------------------------------------

**RelativeAgeUnitID (Primary Key)**

:   An arbitrary Relative Age Unit identification number.

**RelativeAgeUnit**

:   Relative Age Unit. Below are the Relative Age Units for the
    «Geologic time scale» with an example Relative Age.

  --------------------------- ---------------------------
  **Geologic time scale**     

  **RelativeAgeUnit**         **RelativeAge Example**

  Period                      Quaternary

  Epoch                       Pleistocene

  Stage                       Middle Pleistocene

  Informal stage              Middle Holocene
  --------------------------- ---------------------------

«Period», «Epoch», and «Stage» are defined by the International
Commission on Statigraphy. An «Informal stage» is defined in Neotoma.

RelativeChronology {#RelativeChronology}
==================

This table stores relative chronologic data. Relative Ages are assigned
to Analysis Units, The Relative Age data along with any possible
`Geochronology`{.interpreted-text role="ref"} and
`Tephrachronology`{.interpreted-text role="ref"} data are used to create
a chronology.

  --------------------------------------------------------------------------------------
  Field Name                    Variable Type  Key   Reference Table
  ----------------------------- -------------- ----- -----------------------------------
  RelativeChronID               Long Integer   PK     

  AnalysisUnitID                Long Integer   FK    `AnalysisUnits`{.interpreted-text
                                                     role="ref"}

  RelativeAgeID                 Long Integer   FK    `RelativeAges`{.interpreted-text
                                                     role="ref"}

  Notes                         Memo                  
  --------------------------------------------------------------------------------------

**RelativeChronID (Primary Key)**

:   An arbitrary Relative Chronology identification number.

**AnalysisUnitID (Foreign Key)**

:   Analysis Unit identification number. Field links to the
    `AnalysisUnits`{.interpreted-text role="ref"} table.

**RelativeAgeID (Foreign Key)**

:   Relative Age identification number. Field links to the
    `RelativeAges`{.interpreted-text role="ref"} lookup table.

**Notes**

:   Free form notes or comments.

Tephrachronology {#Tephrachronology}
================

This table stores tephrachronologic data. The table relates Analysis
Units with dated tephras in the `Tephras`{.interpreted-text role="ref"}
table. These are tephras with established ages that are used form a
chronology. The tephras are typically not directly dated at the Site of
the Analysis Unit, but have been dated at other sites. A directly dated
tephra, e.g. an argon-argon date, belongs in the
`Geochronology`{.interpreted-text role="ref"} table.

  --------------------------------------------------------------------------------------
  Field Name                   Variable Type  Key    Reference Table
  ---------------------------- -------------- ------ -----------------------------------
  TephrachronID                Long Integer   PK      

  AnalysisUnitID               Long Integer   FK     `AnalysisUnits`{.interpreted-text
                                                     role="ref"}

  TephraID                     Long Integer   FK     `Tephras`{.interpreted-text
                                                     role="ref"}

  Notes                        Memo                   
  --------------------------------------------------------------------------------------

**TephrachronID (Primary Key)** An arbitrary Tephrachronology
identification number.

**AnalysisUnitID (Foreign Key)** Analysis Unit identification number.
Field links to the `AnalysisUnits`{.interpreted-text role="ref"} table.
The tephra may be contained within the AnalysisUnit, especially in
excavations, or the AnalysisUnit may be assigned specifically to the
tephra, particulary with cores.

**TephraID (Foreign Key)** Tephra identification number. Field links to
the `Tephras`{.interpreted-text role="ref"} table.

**Notes** Free form notes or comments about the tephra.

Tephras {#Tephras}
=======

Tephras lookup table. This table stores recognized tephras with
established ages. Referenced by the `Tephrachronology`{.interpreted-text
role="ref"} table.

  ----------------------------------------------
  Field Name             Variable Type    Key
  ---------------------- ---------------- ------
  TephraID               Long Integer     PK

  TephraName             Text             

  C14Age                 Double           

  C14AgeYounger          Double           

  C14AgeOlder            Double           

  CalAge                 Double           

  CalAgeYounger          Double           

  CalAgeOlder            Double           

  Notes                  Memo             
  ----------------------------------------------

**TephraID (Primary Key)**

:   An arbitrary Tephra identification number.

**TephraName**

:   Name of the tephra, e.g. «Mazama».

**C14Age**

:   Age of the tephra in ^14^C yr BP. For example, Hallett et al. (1997)
    provide an estimate of the age of the Mazama tephra based on
    radiocarbon dating of plant macrofossils in lake sediments encasing
    the tephra.

**C14AgeYounger**

:   Younger age estimate of the tephra in ^14^C yr BP.

**C14AgeOlder**

:   Older age estimate of the tephra in ^14^C yr BP.

**CalAge**

:   Age of the tephra in cal yr BP, either calibrated radiocarbon years
    or estimated calendar years derived from another dating method. For
    example, Zdanowicz et al. (1999) identified the Mazama tephra in the
    GISP2 ice core and estimated the age from layer counts.

**CalAgeYounger**

:   Younger age estimate of the tephra in cal yr BP.

**CalAgeOlder**

:   Older age estimate of the tephra in cal yr BP.

**Notes**

:   Free form notes or comments about the tephra.
