# Contact and Individual Related Tables

## `collectors` {#Collectors}

The `collectors` table lists the people who were responsible for collecting a particular collection unit. 

```{r, collectorstable, echo=FALSE}
showTable(db, 'collectors')
```

* **`collectorid (primary key)`**: An arbitrary Collector identification number.
* **`collectionunitid (foreign key)`**]: The `collectionunitid` for the record that was collected. Field links to the [`collectionunits`](#CollectionUnits) table.
* **`contactid (foreign key)`**: Person who collected the CollectionUnit. Multiple individuals are listed in separate records. Field links to the [`contacts`](#Contacts) table.
* **`collectororder`**: Order in which the collectors should be listed.

## `contacts` {#Contacts}

This table lists persons and organizations. The table is referenced
through Foreign Keys in the following tables:
* [`Chronologies`](#Chronologies),
* [`Collectors`](#Collectors),
* [`DatasetPIs`](#DatasetPIs),
* [`DatasetSubmissions`](#DatasetSubmissions),
* [`Projects`](#Projects),
* [`PublicationAuthors`](#PublicationAuthors),
* [`SampleAnalysts`](#SampleAnalysts), and
* [`SiteImages`](#SiteImages) tables.

```{r, contactstable, echo=FALSE}
showTable(db, 'contacts')
```

* **`contactid (primary key)`**: An arbitrary `contactid` number.
* **`aliasid (foreign key)`**: The `contactid` of a person's current name. If the `aliasid` is different from the `contactid`, the `contactid` refers to the person's former name. For example, if J. L. Bouvier became J. B. Kennedy, the `contactid` for J. B. Kennedy is the `aliasid` for J. L. Bouvier.
* **`contactname`**: Full name of the person, last name first (e.g. *Simpson, George Gaylord*) or name of organization or project (e.g. *Great Plains Flora Association*).
* **`contactstatusid (foreign key)`**: Current status of the person, organization, or project. Field links to the [`contactstatuses`](#ContactStatuses) lookup table.
* **`familyname`**: Family or surname name of a person.
* **`leadinginitials**: leading initials for given or forenames **without`** spaces (e.g. *G.G.*).
* **`givennames`**: Given or forenames of a person (e.g. *George Gaylord*). Initials with spaces are used if full given names are not known (e.g. *G. G*).
* **`suffix`**: Suffix of a person's name (e.g. «Jr.», «III»).
* **`title`**: A person's title (e.g. «Dr.», «Prof.», «»).
* **`phone`**: Telephone number.
* **`fax`**: Fax number.
* **`email`**: Email address.
* **`url`**: Universal Resource Locator, an Internet World Wide Web address.
* **`address`**: Full mailing address.
* **`notes`**: Free form notes or comments about the person, organization, or project.

## `contactstatuses` {#ContactStatuses}

Lookup table of Contact Statuses. Table is referenced by the
[`contacts`](#Contacts) table.

```{r, contactstatusestable, echo=FALSE}
showTable(db, 'contactstatuses')
```

* **`contactstatusid (primary key)`**: An arbitrary Contact Status identification number.
* **`contactstatus`**: Status of person, organization, or project.
* **`statusdescription`**:
  * Description of the status. The following statuses exist (with descriptions):
    * **active**: Person, project, or organization is active in the field
    * **deceased**: Person is deceased
    * **defunct**: Project or organization is defunct or non-operational
    * **extant**: Project or organization is extant
    * **inactive**: Person is inactive in the field
    * **retired**: Person is retired
    * **unknown**: Status is unknown
