# Site Related Tables

## SiteImages {#SiteImages}

This table stores hyperlinks to jpeg images of sites.

```{r, siteimagetable, echo=FALSE}
showTable(db, 'siteimages')
```

* **SiteImageID (Primary Key)**: An arbitrary Site Image identification number.
* **SiteID (Foreign Key)**: Site identification number. Field links to the [`Sites`](#Sites) table.
* **ContactID (Foreign Key)**: Contact identification number for image attribution from the [`Contacts`](#Contacts) table.
* **Caption**: Caption for the image.
* **Credit**: Credit for the image. If null, the credit is formed from the    ContactID.
* **Date**: Date of photograph or image.
* **SiteImage**: Hyperlink to a URL for the image.

## Sites {#Sites}

The `sites` table stores information about the geographic information specific to a sampling location, including the name, geographic bounding box, spatial polygon, and site description. Sites generally have an areal extent and can be circumscribed by a latitude-longitude box (using the `latitudenorth` and other numeric bounding variables), but the geospatial object is also stored as a PostGIS `geog` object using a WGS84 projection. Site data ingested from legacy databases may have included only point locations, and additionally, may have been recorded using DMS notation, meaning the decimal expression may appear overly accurate.

The lat-long box can be used either to circumscribe the areal extent of a site or to provide purposeful imprecision to the site location. Site location may be imprecise because of the original description was vague, e.g. «a gravel bar 5 miles east of town», or because the investigators, land owner, or land management agency may not want the exact location made public, perhaps to prevent looting and vandalism. In the first case, the lat-long box can be made sufficiently large to encompass the true location and in the second case to prevent exact location.

```{r, siteimagetable, echo=FALSE}
showTable(db, 'sites')
```

* **SiteID (Primary Key)**: An arbitrary Site identification number.
* **SiteName**: Name of the site.
  * Alternative names, including archaeological site numbers, are placed in square brackets, for example: "New \#4 \[Lloyd\'s Rock Hole\]" and "Modoc Rock Shelter \[11RA501\]".
  * A search of the SiteName field for any of the alternative names or for the archaeological site number will find the site. Some archaeological sites are known only by their site number. Modifiers to site names are placed in parentheses. Authors are added for generic sites names, especially for surface samples, that are duplicated in the database, for example: "Site 1 (Heusser 1978)", "Site 1 (Delcourt et al. 1983)", "Site 1 (Elliot-Fisk et al. 1982)", "Site 1 (Whitehead and Jackson 1990)"
  * Historically, when sitenames were duplicated in the database the name was followed by the 2-letter country code and state or province, for example: "Big Lake (US:Maine)", and "Big Lake (US:Arkansas)". This past practice has declined as the number of database stewards has increased and the capacity to provide improved geospatial searching has improved.
* **LongitudeEast**: East bounding longitude for a site. Currently generated from an internal trigger through PostGIS to extract the bounding box based on the object spatial polygon.
* **LatitudeNorth**: North bounding latitude for a site.
* **LongitudeWest**: West bounding longitude for a site.
* **LatitudeSouth**: South bounding latitude for a site.
* **Altitude**: Altitude of a site in meters.
* **Area**: Area of a site in hectares.
* **SiteDescription**: Free form description of a site, including such information as    physiography and vegetation around the site.
* **Notes**: Free form notes or comments about the site.

## SiteGeoPolitical {#SiteGeoPolitical}

This table lists the GeoPolitical units in which sites occur.

```{r, siteimagetable, echo=FALSE}
showTable(db, 'sitegeopolitical')
```

* **SiteGeoPoliticalID (Primary Key)**: An arbitrary Site GeoPolitical identification number.
* **SiteID (Foreign Key)**: Site identification number. Field links to the [`Sites`](#Sites) table.
* **GeoPoliticalID (Foreign Key)**: GeoPolitical identification number. Field links to the [`GeoPoliticalUnits`](#GeoPoliticalUnits) lookup table.

SQL Example
-----------

The query in Example 2.8.1 lists the GeoPoliticalUnits for «», one unit to a record. This query lists them in a single record.

``` {.sourceCode .sql}
SELECT Sites.SiteName, GeoPoliticalUnits.GeoPoliticalName,
GeoPoliticalUnits\_1.GeoPoliticalName,
GeoPoliticalUnits\_2.GeoPoliticalName

FROM GeoPoliticalUnits AS GeoPoliticalUnits\_2 INNER JOIN
(SiteGeoPolitical AS SiteGeoPolitical\_2 INNER JOIN ((SiteGeoPolitical
AS SiteGeoPolitical\_1 INNER JOIN (GeoPoliticalUnits INNER JOIN (Sites
INNER JOIN SiteGeoPolitical ON Sites.SiteID = SiteGeoPolitical.SiteID)
ON GeoPoliticalUnits.GeoPoliticalID = SiteGeoPolitical.GeoPoliticalID)
ON SiteGeoPolitical\_1.SiteID = Sites.SiteID) INNER JOIN
GeoPoliticalUnits AS GeoPoliticalUnits\_1 ON
SiteGeoPolitical\_1.GeoPoliticalID =
GeoPoliticalUnits\_1.GeoPoliticalID) ON SiteGeoPolitical\_2.SiteID =
Sites.SiteID) ON GeoPoliticalUnits\_2.GeoPoliticalID =
SiteGeoPolitical\_2.GeoPoliticalID

WHERE (((Sites.SiteName)="") AND ((GeoPoliticalUnits.Rank)=1) AND
((GeoPoliticalUnits\_1.Rank)=2) AND ((GeoPoliticalUnits\_2.Rank)=3));
```

Result:

  -------------- ---------------------------------------- ------------------------------------------- -------------------------------------------
  **SiteName**   **GeoPoliticalUnits.GeoPoliticalName**   **GeoPoliticalUnits\_1.GeoPoliticalName**   **GeoPoliticalUnits\_2.GeoPoliticalName**

                                                                                                      Hennepin
  -------------- ---------------------------------------- ------------------------------------------- -------------------------------------------

SQL Example
-----------

The problem with the query above is that if a site has less than three GeoPolitical Names, the result will return empty. For example, «» has no GeoPoliticalUnit with Rank = 3, and will return an empty result with the above query. A solution to this problem is to create and save separate queries for the three ranks:

Query GeoPol1:

``` {.sourceCode .sql}
SELECT Sites.SiteName, GeoPoliticalUnits.GeoPoliticalName

FROM Sites INNER JOIN (GeoPoliticalUnits INNER JOIN SiteGeoPolitical ON
GeoPoliticalUnits.GeoPoliticalID = SiteGeoPolitical.GeoPoliticalID) ON
Sites.SiteID = SiteGeoPolitical.SiteID

WHERE (((GeoPoliticalUnits.Rank)=1));
```

Query GeoPol2:

``` {.sourceCode .sql}
SELECT Sites.SiteName, GeoPoliticalUnits.GeoPoliticalName

FROM Sites INNER JOIN (GeoPoliticalUnits INNER JOIN SiteGeoPolitical ON
GeoPoliticalUnits.GeoPoliticalID = SiteGeoPolitical.GeoPoliticalID) ON
Sites.SiteID = SiteGeoPolitical.SiteID

WHERE (((GeoPoliticalUnits.Rank)=2));
```

Query GeoPol3:

``` {.sourceCode .sql}
SELECT Sites.SiteName, GeoPoliticalUnits.GeoPoliticalName

FROM Sites INNER JOIN (GeoPoliticalUnits INNER JOIN SiteGeoPolitical ON
GeoPoliticalUnits.GeoPoliticalID = SiteGeoPolitical.GeoPoliticalID) ON
Sites.SiteID = SiteGeoPolitical.SiteID

WHERE (((GeoPoliticalUnits.Rank)=3));
```

These three queries can now be combined in a new query with left joins, and the GeoPolitical Names will be returned even if there are less than three.

``` {.sourceCode .sql}
SELECT GeoPol1.SiteName, GeoPol1.GeoPoliticalName,
GeoPol2.GeoPoliticalName, GeoPol3.GeoPoliticalName

FROM (GeoPol1 LEFT JOIN GeoPol2 ON GeoPol1.SiteName = GeoPol2.SiteName)
LEFT JOIN GeoPol3 ON GeoPol2.SiteName = GeoPol3.SiteName

WHERE (((GeoPol1.SiteName)="Lofty "));
```

Result:

  -------------- ------------------------------ ------------------------------ ------------------------------
  **SiteName**   **GeoPol1.GeoPoliticalName**   **GeoPol2.GeoPoliticalName**   **GeoPol3.GeoPoliticalName**

  Lofty                                                                         
  -------------- ------------------------------ ------------------------------ ------------------------------

SQL Example
-----------

The saved queries from the example above can be linked with tables in a more complicated query. This query lists all the pollen sites in the adjacent states of «» in the «» and «» in «».

``` {.sourceCode .sql}
SELECT GeoPol1.SiteName, GeoPol1.GeoPoliticalName,
GeoPol2.GeoPoliticalName, GeoPol3.GeoPoliticalName,
DatasetTypes.DatasetType

FROM (DatasetTypes INNER JOIN ((Sites INNER JOIN CollectionUnits ON
Sites.SiteID = CollectionUnits.SiteID) INNER JOIN Datasets ON
CollectionUnits.CollectionUnitID = Datasets.CollectionUnitID) ON
DatasetTypes.DatasetTypeID = Datasets.DatasetTypeID) INNER JOIN
((GeoPol1 LEFT JOIN GeoPol2 ON GeoPol1.SiteName = GeoPol2.SiteName) LEFT
JOIN GeoPol3 ON GeoPol2.SiteName = GeoPol3.SiteName) ON Sites.SiteName =
GeoPol1.SiteName

GROUP BY GeoPol1.SiteName, GeoPol1.GeoPoliticalName,
GeoPol2.GeoPoliticalName, GeoPol3.GeoPoliticalName,
DatasetTypes.DatasetType

HAVING (((GeoPol1.GeoPoliticalName)="" Or (GeoPol1.GeoPoliticalName)="")
AND ((GeoPol2.GeoPoliticalName)="" Or (GeoPol2.GeoPoliticalName)="") AND
((DatasetTypes.DatasetType)="pollen"))

ORDER BY GeoPol1.GeoPoliticalName, GeoPol2.GeoPoliticalName,
GeoPol3.GeoPoliticalName, DatasetTypes.DatasetType;
```

Result:

  -------------- ------------------------------ ------------------------------ ------------------------------ -----------------
  **SiteName**   **GeoPol1.GeoPoliticalName**   **GeoPol2.GeoPoliticalName**   **GeoPol3.GeoPoliticalName**   **DatasetType**

  Sierra Bacha                                                                                                pollen

  Sierra Bacha 3                                                                                              pollen

                                                                               Apache                         pollen

                                                                               Coconino                       pollen

                                                                               Coconino                       pollen

                                                                               Coconino                       pollen

  Montezuma Well                                                               Yavapai                        pollen
  -------------- ------------------------------ ------------------------------ ------------------------------ -----------------

## GeoPoliticalUnits {#GeoPoliticalUnits}

Lookup table of GeoPoliticalUnits. Table is referenced by the [`SiteGeoPolitical`](#SiteGeoPolitical) table. These are countries and various subdivisions. Countries and subdivisions were acquired from the U.S. Central Intelligence Agency World Factbook[^1] and the ISO 3166-1 and ISO 3166-2 databases[^2].

Each GeoPolitical Unit has a rank. GeoPolitical Units with Rank 1 are generally countries. There are a few exceptions, including Antarctica and island territories, such as , which although a Danish territory, is geographically separate and distinct. Rank 2 units are generally secondary political divisions with various designations: e.g. states in the , provinces in , and regions in . For some countries, the secondary divisions are not political but rather distinct geographic entities, such as islands. The secondary divisions of some island nations include either groups of islands or sections of more highly populated islands; however, the actual island on which a site is located is more important information. Some countries also have Rank 3 units, e.g. counties in the
and metropolitan departments in . In addition to purely political units, various other administrative regions and geographic entities can be
contained in this table. Examples of administrative regions are National Parks and Forests. It might be quite useful, for example, to have a record of all the sites in . These additional units are Rank 4, and they can be added to the database as warranted.

```{r, siteimagetable, echo=FALSE}
showTable(db, 'siteimages')
```

* **GeoPoliticalID (Primary Key)**: An arbitrary GeoPolitical identification number.
* **GeoPoliticalName**: Name of the GeoPolitical Unit, e.g. , .
* **GeoPoliticalUnit**: The name of the unit, e.g. country, state, county, island, governorate, oblast.
* **Rank**: The rank of the unit.
* **HigherGeoPoliticalID**: The GeoPoliticalUnit with higher rank, e.g. the country in which a state lies.

SQL Example
-----------

The following query lists all the political subdivisions of in the .

``` {.sourceCode .sql}
SELECT GeoPoliticalUnits\_2.Rank, GeoPoliticalUnits\_2.GeoPoliticalUnit,
GeoPoliticalUnits\_2.GeoPoliticalName

FROM (GeoPoliticalUnits AS GeoPoliticalUnits\_2 RIGHT JOIN
(GeoPoliticalUnits AS GeoPoliticalUnits\_1 RIGHT JOIN GeoPoliticalUnits
ON GeoPoliticalUnits\_1.HigherGeoPoliticalID =
GeoPoliticalUnits.GeoPoliticalID) ON
GeoPoliticalUnits\_2.HigherGeoPoliticalID =
GeoPoliticalUnits\_1.GeoPoliticalID) LEFT JOIN GeoPoliticalUnits AS
GeoPoliticalUnits\_3 ON GeoPoliticalUnits\_2.GeoPoliticalID =
GeoPoliticalUnits\_3.HigherGeoPoliticalID

WHERE (((GeoPoliticalUnits.GeoPoliticalName)="") AND
((GeoPoliticalUnits\_1.GeoPoliticalName)=""));
```

The first 17 records of the result:

  ------------ ------------------------ ------------------------
  **Rank**     **GeoPoliticalUnit**     **GeoPoliticalName**

  3            district                 Omagh

  3            district                 North Down

  3            district                 Strabane

  3            district                 Newry and Mourne

  3            district                 Moyle

  3            district                 Magherafelt

  3            district                 

  4            historical county        

  4            historical county        

  4            historical county        

  4            historical county        

  4            historical county        

  3            district                 Banbridge

  3            district                 Lisburn

  4            historical county        

  3            district                 Ballymoney

  3            district                 Carrickfergus
  ------------ ------------------------ ------------------------

[^1]: <https://www.cia.gov/library/publications/the-world-factbook/>

[^2]: <http://www.iso.org/iso/country_codes/iso_3166_databases.htm>
